# On map backgrounds {#map}

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```


```{r}
library(broom)
library(dplyr)
library(ggplot2)
library(ggpolypath)
library(cowplot)
library(maps)
library(mapdata)
library(mmap)
library(sp)

theme_map <- function(base_size = 12, base_family = "") {
  theme_bw(base_size = base_size, base_family = base_family) %+replace% 
        theme(axis.title.x = element_text(size = rel(0)),
              axis.title.y = element_text(size = rel(0)))
}
  
theme_set(theme_map())
```

## Objects in the map & mapdata package

Low resolution country coordinates are available in the `map`-packages.

```{r}
m1 <- map("world", region = "Denmark", plot = FALSE, fill = TRUE)
```

The above returns a list object. The `map_data` function in the ggplot2-packages returns a convenient a wrapper of the above which then is "fortified" to a dataframe:

```{r}
m1 <- map_data("world", "Denmark")
```

The function `map_data` always returns the following columns:

* __long__: Longitudinal coordinates
* __lat__:  Latitudinal coordinates
* __group__: A variable, unique to each polygon
* __order__: The plotting order
* __region__: Attribute of the point
* __subregion__: Attribute of the point

Aside from the column containing the coordinates the group variable is the most important. It basically is a unique identifier specifying the polygon to which a particular point belongs to. In order to visualize the data one the following serves as an illustration:
```{r}
p <- ggplot(m1, aes(long, lat)) + coord_quickmap()
p1 <- p + geom_point() + labs(title = "point")
p2 <- p + geom_line() + labs(title = "line")
p3 <- p + geom_path() + labs(title = "path")
p4 <- p + geom_path(aes(group = group)) + labs(title = "path w. group")
plot_grid(p1, p2, p3, p4)
```

Somewhat higher resolution of countries is provided in the `mapdata`-package.
 
```{r}
m2 <- map_data("worldHires", "Denmark") %>% filter(long > 5)
ggplot() +
  aes(long, lat, group = group) +
  geom_path(data = m2, col = "red") +
  annotate("text", x = 13, y = 57.8, label = paste0("high resolution (",nrow(m2)," data points)"), col = "red") +
  geom_path(data = m1, col = "blue") +
   annotate("text", x = 13, y = 58, label = paste0("low resolution (",nrow(m1)," data points)"), col = "blue") +
  coord_quickmap()
```

As seen in the above the difference is quite substantial. In general the resultion needed depends on the objectives (usage).

## ESRI shape files

In the __mmap__-package we have incorporated the ICES ecoregions as a thinned ESRI shapefile object which we will use here as a demonstration:
```{r}
?ecoregion
```

The object attributes are:
```{r}
glimpse(ecoregion@data)
```

Spatial objects can be passed directly to ggplot. What happens under the hood though is that these objects are first converted to a dataframe via the `ggplot2::fortify`-function. This function is being superseeded with the `broom::tidy`-function:
```{r}
glimpse(tidy(ecoregion))
```

Again, here the column of interests for plotting are long, lat and group
```{r}
ggplot() +
  geom_path(data = ecoregion, aes(long, lat, group)) +
  coord_map("ortho", orientation = c(75, -22, 0))
```

As seen above the attributes are not returned when using the `tidy`-function. A simple wrapper to ammend that is provide in the `gisland::tidy2`-function:
```{r}
eco <- gisland::tidy2(ecoregion)
glimpse(eco)
```

```{r}
ggplot() +
  geom_polypath(data = eco, aes(long, lat, group, fill = region)) +
  theme(legend.position = "none") +
  coord_map("ortho", orientation = c(75, -22, 0))
```

Here we use the `ggpolypath::geom_polypath`-function rather than the `ggplot2::geom_polygon`-function because the former deals properly with "holes".

Another example:
```{r}
areas %>% 
  gisland::tidy2() %>% 
  ggplot() +
  geom_polypath(aes(long, lat, group = group, fill = region)) +
  geom_path(aes(long, lat, group = group), lwd = 0.5) +
  theme(legend.position = "none") +
  coord_map("ortho", orientation = c(75, -22, 0))
```

TODO: Think in terms of the chloropleth - in order to save space.

## Backgrounds

```{r, eval = FALSE}
#http://gis.stackexchange.com/questions/76299/web-mapping-service-with-gebco-bathymetry-data
#  http://maps.ngdc.noaa.gov/arcgis/rest/services/web_mercator/gebco08_contours/MapServer/tile/{z}/{y}/{x}
library(leaflet)
leaflet() %>% 
addTiles(urlTemplate = "http://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}",
           attribution = 'Data source: <a href="http://www.hafro.is">Marine Rearch Institute</a>',
           group = "Hnöttur")
leaflet() %>% 
addTiles(urlTemplate = "http://maps.ngdc.noaa.gov/arcgis/rest/services/web_mercator/gebco08_contours/MapServer/tile/{z}/{y}/{x}",
           attribution = 'Data source: <a href="http://www.hafro.is">Marine Rearch Institute</a>',
           group = "Hnöttur")
```


```{r, eval = FALSE}
# Some code on fortify

# http://stackoverflow.com/questions/22096787/how-keep-information-from-shapefile-after-fortify
# 1. approach
# fortify the shape file
map.df <- fortify(shape, region = "CD_GEOCODI")
# merge data
map.df <- left_join(map.df, data, by=c("id" = "CD_GEOCODI"))

# 2. approach
shps <- maptools::readShapePoly(fn = file.path("file",
                                               "path",
                                               "no extension"),
                                     # Column with unique IDs
                                     IDvar = "geoID")
shpsFort <- ggplot2::fortify(model = shps)
```


### Some gisland wrappers

In the __gisland__-package:

```{r, eval = FALSE}
m <- gisland:::map_data2(region = "North Sea")
ggplot() +
  geom_polygon(data = m, aes(long, lat, group = group)) +
  coord_quickmap()
```

## Objects in the mmap package

```{r, eval = FALSE}
m <- mmap::ecoregion[ecoregion$Ecoregion %in% "Iceland Sea",]
ggplot() +
  geom_polypath(data = m, aes(long, lat, group = group)) +
  coord_map()

m <- 
  mmap::ecoregion.thinned %>% 
  filter(Ecoregion %in% "Iceland Sea")
dim(m)
ggplot() +
  geom_polypath(data = m, aes(long, lat, group = group)) +
  coord_map()
```

## google maps and relatives

### ggmap package

```{r}
library(ggmap)
ns <- get_map(location = c(lon = 5, lat = 55),
              zoom = 5,
              maptype = "satellite")
data("harbours", package = "vmstools")
ggmap(ns) +
  geom_point(data = harbours, aes(lon, lat), col = "red", size = 0.25) +
  labs(x = NULL, y = NULL)
```

### Some homebrew

```{r}
library(gisland)
bbox <- c(left = -25.5, bottom = 63, right = -12, top = 67.5)
m <- get_sjokort(bbox = bbox, maptype = "sjm")
m +
  geom_path(data = landhelgi_sldf, aes(long, lat, group = group), col = "red") 
```

```{r}
m <- get_sjokort(bbox <- c(left = -32, bottom = 59.5, right = 12, top = 70),
                 maptype = "norwegian")
m +
  geom_path(data = landhelgi_sldf, aes(long, lat, group = group), col = "red") +
  geom_path(data = eez_sldf, aes(long, lat, group = group), col = "blue") 
```

```{r}
m <- get_sjokort(bbox <- c(left = -32, bottom = 59.5, right = 12, top = 70),
                 maptype = "norwegian")
m +
  geom_path(data = landhelgi_sldf, aes(long, lat, group = group), col = "red") +
  geom_path(data = eez_sldf, aes(long, lat, group = group), col = "blue") 
```

```{r}
maptype = c("World_Imagery", "Ocean_Basemap", 
        "NatGeo_World_Map", "World_Physical_Map", "World_Shaded_Relief", 
        "World_Street_Map", "World_Terrain_Base", "World_Topo_Map")
p <- list()
for (i in 1:length(maptype)) {
  m <- gisland:::get_arcgismap(maptype = maptype[i]) +
    geom_path(data = gisland::landhelgi_sldf, aes(long, lat, group = group)) +
    geom_path(data = gisland::eez_sldf, aes(long, lat, group = group)) +
    labs(title = maptype[i])
  p[[i]] <- m
}

plot_grid(p[[1]], p[[2]], p[[3]], p[[4]],
          p[[5]], p[[6]], p[[7]], p[[8]])
```
