# On gridding data

__Comment__: Need a better title (e.g. summarizing 3d data)

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE)
```

```{r}
library(gisland)
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
```

__A hidden gem?__

`ggplot2::geom_tile`

__Leftover statement__: A very modern example of the (ab)use of such mapping of continuous values is the classification found in the ICES stock summary.)

## On gridding data

```{r, echo = FALSE, eval = FALSE}
# __Generation of example data__:
attach("/home/einarhj/prj/2016_trawlfootprint/data/2013_logbooks_and_vms.rda")
base <- 
  stofn %>% 
  select(tid = visir, vid, gid, towtime = togtimi, t1, t2) %>% 
  filter(tid < 0)  # only electronic logbooks
catch <- 
  afli %>% 
  rename(tid = visir, species = tegund, catch = afli) %>% 
  filter(tid < 0)
tows <-
  base %>% 
  filter(!is.na(t2),
         !is.na(towtime),
         gid %in% 6)
catch2 <- 
  tows %>% 
  select(tid) %>% 
  left_join(catch) %>% 
  group_by(tid) %>% 
  summarise(catch = sum(catch, na.rm = TRUE)) %>% 
  ungroup()
tows <-
  tows %>% 
  left_join(catch2)

vms2 <- 
  vms %>% 
  mutate(tid = lookup_interval_ids(posdate, vid, tows, cn = c("t1", "t2", "vid", "tid"))) %>% 
  filter(!is.na(tid)) %>% 
  select(tid, vid, lon, lat, speed, heading) %>% 
  group_by(tid) %>% 
  mutate(n_pings = n()) %>% 
  ungroup() %>% 
  inner_join(tows %>% select(tid, towtime, catch)) %>% 
  mutate(catch = catch/n_pings,
         effort = towtime/n_pings) %>% 
  mutate(sq = tile_encode(lon, lat, dx = 0.010, dy = 0.005)) %>% 
  group_by(sq) %>% 
  summarise(n = n(),
            effort = sum(effort),
            catch = sum(catch),
            cpue = catch/effort) %>% 
  ungroup() %>% 
  separate(sq, c("lon", "lat"), sep = ":", convert = TRUE)

vms2 %>% 
  filter(between(lon, -25, -24),
         between(lat,  66.5,  67),
         n > 5) %>% 
  ggplot() +
  theme_bw() +
  geom_raster(aes(lon, lat, fill = n)) +
  coord_quickmap() +
  scale_fill_viridis(option = "B", trans = "reverse") +
  labs(x = NULL, y = NULL) +
  theme(legend.position = c(0.8, 0.2))
vms2 %>% 
  filter(between(lon, -25, -24),
         between(lat,  66.5,  67),
         n > 5) %>% 
  ggplot() +
  theme_bw() +
  geom_raster(aes(lon, lat, fill = catch)) +
  coord_quickmap() +
  scale_fill_viridis(option = "B", trans = "reverse") +
  labs(x = NULL, y = NULL) +
  theme(legend.position = c(0.8, 0.2))
vms2 %>% 
  filter(between(lon, -25, -24),
         between(lat,  66.5,  67),
         n > 5) %>% 
  ggplot() +
  theme_bw() +
  geom_raster(aes(lon, lat, fill = cpue)) +
  coord_quickmap() +
  scale_fill_viridis(option = "B", trans = "reverse") +
  labs(x = NULL, y = NULL) +
  theme(legend.position = c(0.8, 0.2))

```

Statistics of two dimentional continuous data (x and y, longitudes and latitudes, ...) are often summarise by mapping the raw data on to a discontinuous grid. The grid is normally a rectangle (tiles) where all the tiles are the same size. A classical case where high resolution data have been reported/summarised to a tiles are the long established catch and effort statistics of the UK fleet, reported by ICES statistical rectangles (the resolution being 1 degree latitude and $1/2$ degree longitude).

```{r, echo = FALSE}
attach("/home/einarhj/stasi/ukSeaCharts.RData")
ukSeaCharts %>% 
  filter(species == 1,
         year == 1932) %>% 
  ggplot() +
  theme_bw() +
  geom_tile(aes(lon, lat, fill = catch)) + 
  facet_wrap(~ month) +
  scale_fill_viridis(option = "B", trans = "reverse") +
  coord_quickmap() +
  scale_x_continuous(NULL, NULL) +
  scale_y_continuous(NULL, NULL) +
  geom_polygon(data = iceland, aes(long, lat, group = group), fill = "grey") +
  labs(title = "Catch distribution of cod by UK fleet by months in 1932")
```

In modern times the "raw" logbook data are normally recorded (at least should) at a much higher resolution. In the Icelandic logbook database the start and end position of each trawl haul has been recorded since 19xx. A generic approach to "round" a continous x- and y-values to a central position is provided by the `gisland::tile_encode` function. Let's take an arbriary example.

```{r}
df.raw <- 
  data_frame(lon = rnorm(2000, -22, 2),
             lat = rnorm(2000, 66, 1))
ggplot(df.raw, aes(lon, lat)) +
  geom_point(alpha = 0.2) +
  coord_quickmap()
```

If we were to count the number of events (points) within a rectangle of same resolution as the ICES rectangle we can do:
```{r}
df <-
  df.raw %>% 
  mutate(sq = tile_encode(x = lon, y = lat, dx = 1, dy = 0.5))
df
```
In the above we have simply "rounded" each record of lon's and the lat's to the defined tile resolution encoded in the character variable `sq`, where the midpoint x- and the y-values are separated by a `:`. To e.g. count the number of records in each tile and return the numerical midpoint values for each tile we simply do:
```{r}
df <- 
  df %>% 
  group_by(sq) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  separate(sq, c("lon", "lat"), sep = ":", convert = TRUE, remove = FALSE)
df
```
The `remove = FALSE` is here simply used for illustrative purpose.

Now to visualize the tile statistics we use the `geom_raster` function:
```{r}
df %>% 
  ggplot(aes(lon, lat)) +
  theme_bw() +
  geom_raster(aes(fill = n)) +
  geom_point(data = df.raw, aes(lon, lat), colour = "red", size = 0.25) +
  geom_text(aes(label = n), colour = "yellow", angle = 45) +
  coord_quickmap()
```


An attempt to provide generic approach is wrapped in the `gisland::tile_encode` function:

```{r, message = FALSE}
attach("/home/einarhj/prj/2016_trawlfootprint/data/2013_logbooks_and_vms.rda")
base <- stofn
tows <-
  base %>% 
  select(t1, t2, vid, gid, visir) %>% 
  filter(!is.na(t2),
         gid %in% 6)
vms2 <- 
  vms %>% 
  mutate(visir = lookup_interval_ids(posdate, vid, tows, cn = c("t1", "t2", "vid", "visir"))) %>% 
  filter(!is.na(visir)) %>% 
  mutate(sq = tile_encode(lon, lat, dx = 0.010, dy = 0.005)) %>% 
  group_by(sq) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  separate(sq, c("lon", "lat"), sep = ":", convert = TRUE)

vms2 %>% 
  filter(between(lon, -25, -24),
         between(lat,  66.5,  67),
         n > 5) %>% 
  ggplot() +
  theme_bw() +
  geom_raster(aes(lon, lat, fill = n)) +
  coord_quickmap() +
  scale_fill_viridis(option = "B", trans = "reverse") +
  labs(x = NULL, y = NULL) +
  theme(legend.position = c(0.8, 0.2))

# zoom out a bit
vms2 %>% 
  filter(between(lon, -27, -23),
         between(lat,  65.75,  67.3),
         n > 5) %>% 
  ggplot() +
  theme_bw() +
  geom_polygon(data = iceland, aes(long, lat, group = group), fill = "grey") +
  geom_raster(aes(lon, lat, fill = n)) +
  coord_quickmap(xlim = c(-27, -23), ylim = c(65.75,  67.3)) +
  scale_fill_viridis(option = "B", trans = "reverse") +
  labs(x = NULL, y = NULL) +
  theme(legend.position = c(0.1, 0.8))

# and then some
vms2 %>% 
  filter(between(lon, -27, -9),
         between(lat,  62.8,  67.75)) %>% 
  mutate(n = ifelse(n > 50, 50, n)) %>% 
  ggplot() +
  theme_bw() +
  geom_polygon(data = iceland, aes(long, lat, group = group), fill = "grey") +
  geom_raster(aes(lon, lat, fill = n), interpolate = TRUE) +
  coord_quickmap() +
  scale_fill_viridis(option = "B", trans = "reverse") +
  labs(x = NULL, y = NULL) +
  theme(legend.position = c(0.5, 0.4))
```


## ICES rectangles

A historical case, that has been with us in fishiers statistics (far to long) is to map catches from a particular fishing operation on statistical squares. Be they ices statistical rectangles (with a resolution 1° longitude and 0.5° latitude) or any other native tile format. Just as a case in point that this has been with us for a long time, here are some some data of the English fleet catch by ICES rectangle from the 19xx:

```{r}


```

In the above case, as in any other rectangle/tile code system one needs to decode a one dimentional vector (normally a character) into some (dis-)continous values that are suitable for plotting. In the above case this is done with the `decode_tile`-function.

## c-square

Another example is the format that ICES report to OSPAR.

... something more on c-square ...


```{r}


```

... the problem is the same, to decode one dimentional character vector into x- and y-dimensions so it can be plotted. In the above case, we use the same `decode_tile`-function but specify instead of using the default (system = "ICES") we use (system = "csquare").




```{r}
data(tacsat, package = "vmstools")
tacsat %>% 
  gisland::tidy_tacsat() %>% 
  select(SI_LONG, SI_LATI) %>% 
  mutate(sq = tile_encode(SI_LONG, SI_LATI, dx = 0.50, dy = 0.25)) %>% 
  group_by(sq) %>% 
  slice(1)
```

The function returns the midpoints of a tile, whose resolution is specifed by _dx_ and _dy_ as a character vector where the midpoints are parsed, separated by _:_. Based on this one can start to create various summary statistics on each tile, e.g.

```{r}
vms <- 
  tacsat %>% 
  tidy_tacsat() %>% 
  select(year, SI_LATI, SI_LONG, SI_SP, SI_HE) %>% 
  mutate(sq = tile_encode(SI_LONG, SI_LATI, dx = 0.50, dy = 0.25)) %>% 
  group_by(year, sq) %>% 
  summarise(n = n(),
            # not necessarily sensible statistics
            maxSP = max(SI_SP, na.rm = TRUE),
            minSP = min(SI_SP, na.rm = TRUE),
            meanSP = mean(SI_SP, na.rm = TRUE),
            medianHE = median(SI_HE, na.rm = TRUE))
glimpse(vms)
```

To get back the numerical value of the tile midpoints one can use the `tidyr::separate` function.
```{r}
vms <- 
  vms %>% 
  separate(sq, c("lon", "lat"), sep = ":", convert = TRUE)
glimpse(vms)
```

A visual map can then be obtained via the `ggplot2::geom_tile` function:

```{r}
library(mmap)
m <- ecoregion[ecoregion$Ecoregion %in% "Greater North Sea",]
vms %>% 
  filter(between(lon, -4, 10),
         between(lat, 25, 58)) %>% 
  ggplot() +
  geom_polygon(data = m, aes(long, lat, group = group), fill = "white") +
  geom_tile(aes(lon, lat, fill = n)) +
  scale_fill_viridis(option = "B", trans = "reverse") +
  facet_wrap(~ year) +
  coord_quickmap()
rm(vms)
```

