# On point attributes

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE)
```

```{r}
library(sp)
library(gisland)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpolypath)

theme_map <- function(base_size = 12, base_family = "") {
  theme_bw(base_size = base_size, base_family = base_family) %+replace% 
        theme(axis.title.x = element_text(size = rel(0)),
              axis.title.y = element_text(size = rel(0)))
}
  
theme_set(theme_map())
```

## Finding point attributes

When working with spatial data one is often interested in finding the attributes of a particular points/coordinates. Like if a point falls within a certain area, region, strata, ... etc. If the data are of the sp-class family one can generally solve this using the `sp::over`-function. This function has been wrapped inside the `gisland::geo_inside`-function so that it falls under the tidyverse pipe flow.

Generally we would have:

* A vector of longitude and latitude coordinates
* A spatialPolygonDataFrame with some attributes

What we would like as the output is:

* A vector of same length as the points that represents the region that the coordinates correpond to.

### Examples

#### ICES area

```{r}
map <- mmap::areas[mmap::areas$region %in% c("Va1", "Va2"),]
class(map)
```

The attributes are:
```{r}
map@data
```

Lets generate some fictionary spatial positions:
```{r}
set.seed(16)
station <- 
  data_frame(long = runif(1000, -28, -9),
             lat  = runif(1000, 61.5, 68.5))
```

If we were interested in obtaining the "ICES_area" to which the points belong to one could do:
```{r}
st <-
  station %>% 
  mutate(strata = geo_inside(long, lat, map, variable = "region"))
glimpse(st)
st %>% 
  group_by(strata) %>% 
  summarise(n = n())
```

So here we have `r length(st$strata[st$strata %in% "Va1"])` points that fall within strata Va1, `r length(st$strata[st$strata %in% "Va2"])` points that fall within strata Va2 and `r sum(is.na(st$strata))` points that are outside those two stratas. A visual representation can be obtained as:
```{r}
ggplot(map, aes(x = long, y = lat)) +
  geom_polypath(aes(group = group), fill = "grey") +
  geom_path(aes(group = group), col = "black") +
  geom_point(data = st, aes(col = strata), size = 1) +
  coord_quickmap()
```

Take note that points that are on land (within the Icelandic shoreline) are not included within strata Va2. This is because the shorelines are defined as "holes" within the Va2 object.

If one were only interested in subsetting the data set, keeping only the points that fall within an area one could simply not specify the variable name in the `geo_inside`-function. The output would then simply be a boolean vector which one could pass to the `dplyr::filter`-function.

```{r}
station %>%  
  filter(geo_inside(long, lat, map)) %>% 
  nrow()
```

#### Points on land

This is kind of a reverse problem from the one above. In order to classify if coordinates fall on land one could do:
```{r}
st <- 
  station %>% 
  mutate(on_land = geo_inside(long, lat, biceland))
st %>% 
  group_by(on_land) %>% 
  summarise(n = n())
```

Visually we have:
```{r}
st %>% 
  ggplot() +
  geom_polygon(data = biceland, aes(long, lat, group = group), fill = "grey") +
  geom_point(aes(long, lat, colour = on_land), size = 1) +
  coord_quickmap()
```

One could think that one of the point on the map (the southernmost on_land point) has been wrongly classified. That is however not the case because it falls within one of the small islands of Vestmannaeyjar.

If one were to exclude the points on land one would use `filter(!on_land)` in the pipe above.

#### Points in vicinity of some coordinates

We may be interested in excluding points that are in vicinity of some coordinates, like points in the vicinity of harbour. If coordinates are only given up as a points, like in a data.frame we need first somehow to convert the data to spatial objects. This can be achieved using the `gisland::expand_sp`-function (a simple wrapper around the `rgeos::gBuffer`-function).

The process as follows:

* We have a data.frame of interest
* Generate a spatial point dataframe
* Generate a spatial polygon were each of the point is "expanded" to a circle by specifying some radius (here 10 miles)
* Identify the points that fall within the spacial polygon (vicinity to e.g. harbour)


```{r}
df <- data.frame(long = c(-20, -15),
                 lat = c(64, 64))
sp <- 
  SpatialPointsDataFrame(df, df, proj4string = PRO) %>% 
  expand_sp(miles = 20)
# note would really like to generate a sp-dataframe so we also have attributes

st <- 
  station %>% 
  filter(!geo_inside(long, lat, iceland)) %>% 
  mutate(in_harbour = geo_inside(long, lat, sp)) 
st %>% 
  ggplot(aes(long, lat)) +
  geom_polypath(data = iceland, aes(group = group), fill = "grey90") +
  geom_path(data = sp, aes(group = group)) +
  geom_point(aes(colour = in_harbour)) +
  coord_quickmap()
```

The `gisland::expand_sp`-function can be used on spatitalPolygons. E.g. if we were interested in locating points that are within 12 mile vicinity of the icelandic coast we could:
```{r}
sp <- expand_sp(iceland, miles = 12) # takes a while
st <-
  station %>% 
  filter(!geo_inside(long, lat, iceland)) %>% 
  mutate(within_3_miles = geo_inside(long, lat, sp))

st %>% 
  ggplot(aes(long, lat)) +
  geom_polypath(data = iceland, aes(group = group), fill = "grey90") +
  geom_path(data = sp, aes(group = group)) +
  geom_point(aes(colour = within_3_miles)) +
  coord_quickmap()
```

As seen in the above cases the procedures of finding coordinate attributes is always the same:

* pass vector of coordinates and a spatial polygon to the `geo_inside` function.
* if needed, generate a spatial polygon or modify existing one with the help of the `expand_sp` function.

The generation of spatial objects is an art in itself, something that will not be dealt more than the limited examples provided above.

