# Appendicies

## Preamble

Here I describe how the sqlite database was populated.

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE)
```

```{r, message = FALSE}
library(dplyr)
library(ggplot2)
library(ggpolypath)
```

## Ecoregion

```{r, eval = FALSE}
mmap <- src_sqlite("/home/einarhj/r/Pakkar/mmap/inst/extdata/mmap.sqlite3")
ecoregion <- readOGR(dsn = "/home/einarhj/stasi/gis/ices/data-raw/ICES_ecoregions/",
               layer = "ICES_ecoregions_20150113_no_land")

ecoregiondat <- 
  ecoregion@data %>% 
  mutate(object_ = row_number()) %>% 
  tbl_df()
ecoregion <- spbabel::sptable(ecoregion)
copy_to(mmap, ecoregion, temporary = FALSE, indexes = list(c("object_", "branch_"), c("x_", "y_")))
copy_to(mmap, ecoregiondat, temporary = FALSE, indexes = list("Ecoregion"))
```

### Usage

#### A function

```{r}
get_ecoregion <- function(region = c("Greenland Sea", "Iceland Sea")) {
  
  path <- system.file("extdata/mmap.sqlite3", package = "mmap")
  mmap <- dplyr::src_sqlite(path)
  
  d <- 
    dplyr::tbl(mmap, "ecoregiondat") %>% 
    dplyr::filter(Ecoregion %in% region) %>% 
    dplyr::inner_join(tbl(mmap, "ecoregion"), by = "object_") %>% 
    dplyr::collect() %>% 
    dplyr::select(region = Ecoregion, long = x_, lat = y_, group = branch_)
  
  return(d)
  
}
```

#### Example

```{r}
m <- get_ecoregion()

ggplot() +
  geom_polypath(data = m, aes(long, lat, group = group, fill = region)) +
  coord_quickmap()
```

## Depths (Sjómælingar)

The raw data were obtained from Icelandic Coastguard - Hydrographic Department via Julian Burgos through an e-mail dated 2015-06-29:
```
Hi Einar,
I haven't had the time to add them to the gisland package.  But here are
the shapefiles.  When I asked Árni Vésteinsson about the sources of the
data, he answered this:
"The contours come from our Nautical Charts at scale 1:100.000 and smaller
(300.000, 1:1000.000).
Our Chart Catalogue is here:
http://www.lhg.is/starfsemi/sjomaelingasvid/sjokort/kortaskra/
The charts that I am refering to are the Coastal and Transit charts, page
6 and 8.
The data source is mainly ICG Hydrographic Department. There are other
sources too but I suppose you don´t want all the details."
```

The raw data are found within the `data_sp_raw`-directory:
```{r, echo = FALSE}
dir("/home/einarhj/stasi/gis/sjomaelingar/data_sp_raw")
```

### Processing

The main objective was to create spatial polygons from the spatial lines objects.

#### Cleaning the spatial lines objects

The original shape data in `data_sp_raw/Isl_dyp_lin_290115` are spatial lines. Each depth was separated into a separate object using the script 'R/01_separated_z.R` and saved in directory `data_spatial_lines`:

```{r, echo = FALSE, message = FALSE}
library(dplyr)
library(stringr)
dir("/home/einarhj/stasi/gis/sjomaelingar/data_spatial_lines", pattern = ".dbf") %>%
  str_replace(".dbf","")
```

Some line objects needed to be cleaned:

* Line segments in some areas, particularily in Ísafjarðardjúp were joined.
* Closing of some line objects to form a "polygon"
* Cleaning of some line objects because of intersection problems.

This process was done manually in qgis. The cleanded objects were saved under the same file name in the `data_spatial_lines`-directory (see above).

#### Creation of spatial polygons

Spatial polygons objects were generated from spatial lines objects using the following scripts:
```{r, echo = FALSE}
dir("/home/einarhj/stasi/gis/sjomaelingar/R", pattern = "02")
```


### Output examples

See:

* [leaflet map](sjomaelingar.html) - generated via `R/03leaflet.R`
* qgis file `sjomaelingar.qgs` in the root of the working directory

#### To do

- Put the separate depth polygons files into one file

#### Loading

```{r, eval = FALSE}
library(ggplot2)
library(ggpolypath)
library(rgdal)
files <- dir("/home/einarhj/stasi/gis/sjomaelingar/data_spatial_lines",
             pattern = ".shp") %>% 
  str_replace(".shp","")

Z <- readOGR(dsn = "/home/einarhj/stasi/gis/sjomaelingar/data_spatial_lines/",
             layer = files[1]) %>% 
  spTransform(gisland::PRO)
Z <-  spChFIDs(Z, paste(row.names(Z), Z$MinZ, sep = "_"))

for (i in 2:length(files)) {
  z <- readOGR(dsn = "/home/einarhj/stasi/gis/sjomaelingar/data_spatial_lines/",
             layer = files[i]) %>% 
  spTransform(gisland::PRO)
  z <-  spChFIDs(z, paste(row.names(z), z$MinZ, sep = "_"))
  #print(paste(i, files[i]))
  #print(row.names(z)[1])
  Z <- spRbind(Z, z)
}

depth <- 
  Z %>% 
  gisland::fortify2() %>% 
  select(z = Minz, group, long = lon, lat)
#mmap$con %>% db_drop_table(table='depth')

mmap <- src_sqlite("/home/einarhj/r/Pakkar/mmap/inst/extdata/mmap.sqlite3")

# check the correct order
copy_to(mmap,
        depth,
        temporary = FALSE)
```

Alternatives (not working):
```{r, eval = FALSE}
src_sqlite("test.sqlite", create = TRUE)
writeOGR(Z, "/home/einarhj/prj/fishy-data/test.sqlite",layer = "Z", driver = "SQLite",
         dataset_options=c("SPATIALITE=yes"))
writeOGR(Z, dsn="./test.sqlite",layer="test",driver= "SQLite",overwrite_layer=TRUE)
writeOGR(Z, dsn="./test.sqlite",layer="test",driver= "SQLite")
writeOGR(Z, dsn="/home/einarhj/prj/fishy-data/test.sqlite",layer="test",driver= "SQLite",overwrite_layer=TRUE)
writeOGR(Z, dsn="/home/einarhj/prj/fishy-data/test.sqlite",layer="test",driver= "SQLite")
```

#### Example

```{r}
library(dplyr)
library(ggplot2)

mmap <- src_sqlite("/home/einarhj/r/Pakkar/mmap/inst/extdata/mmap.sqlite3")
depth <- tbl(mmap, "depth")

d2 <-
  depth %>% 
  filter(between(long, -24, -22),
         between(lat, 65, 65.5)) %>% 
  collect()

ggplot() +
  theme_bw() +
  geom_path(data = d2, aes(long, lat, group = group, colour = factor(z)), lwd = 0.3) +
  scale_colour_brewer(palette = "Set1") +
  geom_polygon(data = gisland::iceland, aes(long, lat, group = group), fill = "grey90") +
  coord_quickmap(xlim = c(-24, -22), ylim = c(65, 65.5)) +
  labs(x = NULL, y = NULL, colour = "depth")
```

